Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-04 PandaEvolution.tex
Notes clean-up

The next LHC run for taking data (Run 3) will require more resources than the
Worldwide LHC Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more
than
% 100,000
300,000 cores at over 100 Grid sites\mtnote{Addresses Alexei's comment.
Increased Grid capacity to 300K cores. Is the number of sites still 100 or
should I increase that too?}, with a peak performance of 0.3 petaFLOPS. This
capacity will be sufficient for the planned analysis and data processing, but it
will be insufficient for the Monte Carlo production workflow and any extra
activity. To alleviate these challenges, ATLAS is engaged in a program to expand
the current computing model to include additional resources such as the
opportunistic use of supercomputers.
% as well as commercial and academic clouds.

@@

\mtnote{Move this to the introduction: ``There are at least two approaches to
enable PanDA WMS to execute ATLAS workloads or workflows on supercomputers: (i)
using the subset of resources and capabilities shared by both supercomputers and
WLCG; (ii) reconciling the parallel and distributed computing paradigms by means
of dedicated abstractions. The former is a pragmatic approach that enables the
execution of specific workloads by prototyping single-point solutions. The
latter is a principled approach, better suited for a production-grade solution,
capable of supporting general-purpose workloads and workflows on both
supercomputers and Grid infrastructures. \\ This section illustrates the design
and architecture of a job broker prototyped by the PanDA team. This broker
supports execution of part of the ATLAS production Monte Carlo workflow on
Titan, a leadership-class supercomputer managed by the Oak Ridge Leadership
Computing Facility (OLCF) at the Oak Ridge National Laboratory (ORNL).After an
analysis of the results obtained and the lessons learned, the following section
introduces the design and first experimental characterization of a next
generation executor (NGE). NGE is designed to abstract resources and
capabilities, enabling the concurrent execution of both parallel and distributed
computing on generic high performance computing (HPC) machines.''}

\jhanote{Do you mean the top of section 3? Or do you mean the very beginning
section?}

@@

Detector simulation is part of the ATLAS production Monte Carlo (MC) workflow
(also known as MC production
chain)~\cite{rimoldi2006atlas,de2013delphes,ritsch2014atlas}. The MC workflow
consists of four main stages: event generation, detector simulation,
digitization, and reconstruction. Event generation creates sets of particle
four-momenta via different generators, e.g., PYTHIA~\cite{sjostrand2006pythia},
HERWIG~\cite{corcella2001herwig} and many others. \mtnote{Addresses Alexei's
comment. Is the following sentence correct?} The detector simulator is called
Geant4~\cite{agostinelli2003geant4} and simulates the ATLAS detector and the
interaction between that detector and particles.
% the interaction of these particles with the sensitive material of the ATLAS
% detector.
Each interaction creates a so-called hit and all hits are collected and passed
on for digitalization, where hits are further processed to mimic the readout of
the detector. Finally, reconstruction operates local pattern recognition,
creating high-level objects like particles and jets.

@@

Backfill querying is performed via a dedicated Moab scheduler command while a
tailored Python MPI script is used to execute the payload of ATLAS jobs. This
MPI script enables the execution of unmodified Grid-centric, ATLAS jobs on
Titan. Typically, a MPI script is workload-specific as it sets up the execution
environment for a specific payload. This involves organization of worker
directories, data management, optional input parameters modification, and
cleanup on exit. Upon submission, a copy of the MPI script runs on every
available worker node,
% Each script knows its MPI rank (an index that runs from zero to a maximum
% number of nodes or script copies) as well as the total number of ranks in the
% current submission. When activated on worker nodes, each copy of the MPI
% script
starting the execution of the ATLAS job's payload in a subprocess and waits
until its completion.

@@

\mtnote{Address Alexei's comments: Add a paragraph explaining how this solution
has been ported to other HPC machines and reference relevant publications.}
While PanDA Broker implementation is resource specific, it was successfully
ported also to other supercomputers, including the Ansel (?) at \ldots and
Cori at the National Energy Research Scientific Computing Center (NERSC).

>>>>

The next LHC run for taking data (Run 3) will require more resources than the
Worldwide LHC Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more
than 300,000 cores at over 100 Grid sites, with a peak performance of 0.3
petaFLOPS. This capacity will be sufficient for the planned analysis and data
processing, but it will be insufficient for the Monte Carlo production workflow
and any extra activity. To alleviate these challenges, ATLAS is engaged in a
program to expand the current computing model to include additional resources
such as the opportunistic use of supercomputers.

% as well as commercial and academic clouds.
% \mtnote{Addresses Alexei's comment. Increased Grid capacity to 300K cores. Is
% the number of sites still 100 or should I increase that too?}

@@

% \mtnote{Move this to the introduction: ``There are at least two approaches to
% enable PanDA WMS to execute ATLAS workloads or workflows on supercomputers: (i)
% using the subset of resources and capabilities shared by both supercomputers and
% WLCG; (ii) reconciling the parallel and distributed computing paradigms by means
% of dedicated abstractions. The former is a pragmatic approach that enables the
% execution of specific workloads by prototyping single-point solutions. The
% latter is a principled approach, better suited for a production-grade solution,
% capable of supporting general-purpose workloads and workflows on both
% supercomputers and Grid infrastructures. \\ This section illustrates the design
% and architecture of a job broker prototyped by the PanDA team. This broker
% supports execution of part of the ATLAS production Monte Carlo workflow on
% Titan, a leadership-class supercomputer managed by the Oak Ridge Leadership
% Computing Facility (OLCF) at the Oak Ridge National Laboratory (ORNL).After an
% analysis of the results obtained and the lessons learned, the following section
% introduces the design and first experimental characterization of a next
% generation executor (NGE). NGE is designed to abstract resources and
% capabilities, enabling the concurrent execution of both parallel and distributed
% computing on generic high performance computing (HPC) machines.''}
%
% \jhanote{Do you mean the top of section 3? Or do you mean the very beginning
% section?}

@@

% \mtnote{Addresses Alexei's comment. Is the following sentence correct?}

Detector simulation is part of the ATLAS production Monte Carlo (MC) workflow
(also known as MC production
chain)~\cite{rimoldi2006atlas,de2013delphes,ritsch2014atlas}. The MC workflow
consists of four main stages: event generation, detector simulation,
digitization, and reconstruction. Event generation creates sets of particle
four-momenta via different generators, e.g., PYTHIA~\cite{sjostrand2006pythia},
HERWIG~\cite{corcella2001herwig} and many others. The detector simulator is
called Geant4~\cite{agostinelli2003geant4} and simulates the ATLAS detector and
the interaction between that detector and particles. Each interaction creates a
so-called hit and all hits are collected and passed on for digitalization, where
hits are further processed to mimic the readout of the detector. Finally,
reconstruction operates local pattern recognition, creating high-level objects
like particles and jets.

@@

Backfill querying is performed via a dedicated Moab scheduler command while a
tailored Python MPI script is used to execute the payload of ATLAS jobs. This
MPI script enables the execution of unmodified Grid-centric, ATLAS jobs on
Titan. Typically, a MPI script is workload-specific as it sets up the execution
environment for a specific payload. This involves organization of worker
directories, data management, optional input parameters modification, and
cleanup on exit. Upon submission, a copy of the MPI script runs on every
available worker node, starting the execution of the ATLAS job's payload in a
subprocess and waits until its completion.

% Each script knows its MPI rank (an index that runs from zero to a maximum
% number of nodes or script copies) as well as the total number of ranks in the
% current submission. When activated on worker nodes, each copy of the MPI
% script

@@

% \mtnote{Address Alexei's comments: Add a paragraph explaining how this solution
% has been ported to other HPC machines and reference relevant publications.}
While PanDA Broker implementation is resource specific, it was successfully
ported also to other supercomputers, including the Ansel (?) at \ldots and
Cori at the National Energy Research Scientific Computing Center (NERSC).
